\begin{enumerate}[label=\thechapter.\arabic*,ref=\thechapter.\theenumi]
\item Let $\phi(.)$ denote the cumulative distribution function of a standard normal
random variable. If the random variable $X$ has the cumulative distribution
function 
\begin{align}
	F(x)&= 
    \begin{cases}
        \phi(x), &  x < -1 \\
        \phi(x+1) , &  x \ge -1
    \end{cases} 
\end{align}
then which one of the following statements is true?
\begin{enumerate}
\item $P(X \leq -1) = \frac{1}{2}$
\item $P(X = -1) = \frac{1}{2}$
\item $P(X < -1) = \frac{1}{2}$
\item $P(X \leq 0) = \frac{1}{2}$
\end{enumerate}
\hfill(GATE ST 2023)
\\
\input{2023/ST/15/main.tex}
\item Let $X$ be a random variable with the probability density function $f(x)$ such that
\begin{align}
f(x) &= 
	\begin{cases}
		\frac{1}{2\sqrt{3}}, & -\sqrt{3} \leq x \leq \sqrt{3} \\
		0, & \text{otherwise}
	\end{cases}
\end{align}
Then the variance of $X$ is?
\hfill(GATE XH-C1 2023)
\\
\input{2023/XH/62/main.tex}
\item Two defective bulbs are present in a set of five bulbs. To remove the two
defective bulbs, the bulbs are chosen randomly one by one and tested. If $X$
denotes the minimum number of bulbs that must be tested to find out the two
defective bulbs, then $\pr{X=3}$ (rounded off to two decimal places)
equals\\
\hfill(GATE ST 2023)\\
\input{2023/ST/31/main.tex}
\item Let $X$ be a random variable with cumulative distribution function
\begin{align}
F_X(x) &= 
	\begin{cases}
		0 & \text{if $x < -1$}\\
		\frac{1}{4}\brak{x+1} & \text{if $-1 \leq x < 0$}\\
		\frac{1}{4}\brak{x+3} & \text{if $0 \leq x < 1$}\\
		1 & \text{if $ x \geq 1$}\\
	\end{cases}
\end{align}
Which one of the following statements is true?
\begin{enumerate}[label=(\Alph*)]
    \item\begin{align} \lim_{n \to \infty} \pr{-\frac{1}{2} + \frac{1}{n} < X < -\frac{1}{n}} = \frac{5}{8}\end{align}
    \item \begin{align}\lim_{n \to \infty} \pr{-\frac{1}{2} - \frac{1}{n} < X < \frac{1}{n}} = \frac{5}{8}\end{align}
    \item \begin{align}\lim_{n \to \infty} \pr{X = \frac{1}{n}} = \frac{1}{2}\end{align}
    \item \begin{align}\pr{X = 0} = \frac{1}{3}\end{align}
\end{enumerate}

\hfill(GATE ST 2023)\\
\input{2023/ST/46/main.tex}
\item Three unbiased coins were tossed. Provided that at least two outcomes are tails,the probability of having all three outcomes as tails is\\
\hfill(GATE PI 2023)\\
\input{2023/PI/14/main.tex}
\item Let $X$ be a random variable with probability density function
\begin{align}
p_{X}\brak{x}=\begin{cases}
		e^{-x} & if x\ge 0\\ 
		0 & otherwise
	\end{cases}
\end{align}
For $a<b$, if $U\brak{a,b}$ denotes the uniform distribution over the interval $\brak{a,b}$,
then which of the following statements is/are true?
\begin{enumerate}[label=(\Alph*)]
\item $e^{-X}$ follows $U\brak{-1,0}$ distribution
\item $1-e^{-X}$follows $U\brak{0,2}$ distribution
\item $2e^{-X}-1$follows $U\brak{-1,1}$ distribution
\item The probability mass function of $Y=\sbrak{X}$ is
$\pr{Y=k}=e^{-k}\brak{1-e^{-1}}$ for k= 0, 1, 2, …,
where $\sbrak{X}$denotes the largest integer not exceeding $x$
\end{enumerate}
\hfill(GATE ST 2023)\\
\solution
\input{2023/ST/53/main.tex}
\item Question: Let $X$ be a positive valued continuous random variable with finite mean $\mu$.
If $Y=[X]$, the largest integer less than or equal to $X$, then which of the
following statements is/are true?
\begin{enumerate}[label=(\Alph*)]
\item $\pr{Y \leq \mu} \leq \pr{X \leq \mu}$ for all $\mu \geq 0$
\item $\pr{Y \geq \mu} \leq \pr{X \geq \mu}$ for all $\mu \geq 0$
\item E(X) $<$ E(Y)
\item E(X) $>$ E(Y)
\end{enumerate}
\hfill(GATE ST 2023)\\
\input{2023/ST/52/main.tex}
\item In a diploid angiosperm species, flower colour is regulated by the R gene.
RR and Rr genotypes produce red flowers, whereas the rr genotype produces
white flowers. If two individual plants are randomly selected from a large
segregating population of a genetic cross between RR and rr parents, the
probability of both the plants producing red flowers is\\
\hfill(GATE XL 2023)\\
\input{2023/XL/63/main.tex}
\item The frequencies for autosomal alleles $A$ and $a$ are $p = 0.5$ and $q = 0.5$,
respectively, where $A$ is dominant over $a$. Under the assumption of random
mating, the mating frequency among dominant parents is.\\
\hfill(GATE XL 2023)\\
\input{2023/XL/65/main.tex}
\item Let $X$ be a random variable having poisson distribution with mean $\lambda>0$. Then E\brak{\cond{\frac{1}{1+X}}{X>0}} equals
\begin{enumerate}
	\item $\frac{1-e^{-\lambda}-\lambda e^{-\lambda}}{\lambda\brak{1-e^{-\lambda}}}$
	\item $\frac{1-e^{-\lambda}}{\lambda}$
	\item $\frac{1-e^{-\lambda}-\lambda e^{-\lambda}}{\lambda}$
	\item $\frac{1-e^{-\lambda}}{\lambda +1}$
\end{enumerate}
\hfill(GATE ST 2023)\\
\solution\\
\input{2023/ST/17/main.tex}
\item Consider the probability space $(\Omega, \mathcal{G}, P)$, where 
   $ \Omega = \{1, 2, 3, 4\}$, 
    $\mathcal{G} = \{\emptyset, \Omega, \{1\}, \{4\}, \{2, 3\}, \{1, 4\}, \{1, 2, 3\}, \{2, 3, 4\}\}$, 
    $P(\{1\}) = \frac{1}{4}$.
Let $X$ be the random variable defined on the above probability space as
   $ X(1) = 1$, 
    $X(2) = X(3) = 2$, 
    $X(4) = 3$.
If $P(X \leq 2) = \frac{3}{4}$, then find $P(\{1, 4\})$ (rounded off to two decimal places).\\\hfill (GATE ST 2023)\\
\input{2023/ST/60/main.tex}
Let ${N(t)}_{t\ge 0}$ be a Poisson process with rate 1. Consider the following statements. 
\begin{enumerate}[label=(\alph*)]
\item $\pr{N(3)=3|N(5)=5}=\comb{5}{3}\left(\frac{3}{5}\right)^3 \left(\frac{2}{5}\right)^2$
\item If $S_5$ denotes the time of occurrence of the $5^{th}$ event for the above Poisson process,then $E(S_5|N(5)=3)=7$ \\
Which of the above statements is/are true?\\
\end{enumerate}
\begin{enumerate}[label=(\roman*)]
\item only (a)
\item only (b)
\item Both (a) and (b)
\item Neither (a) and (b)
\end{enumerate}
\hfill ( GATE ST 2023)\\
\solution \\
\input{2023/ST/38/main.tex}
\end{enumerate}
